# Completeness checks (Dorian & Cassie)


We should check for data completeness before generating summaries for sampling data. 

Completeness checks inlcude tests for:  
  
- Seasonal completeness (75%)
    - This is an EPA guideline for criteria pollutants and has been applied to air toxics (citation).
- 20% values above detection
    - This is based on _Cox 2006_ and MPCA simulations (cite work and results)
- Label for criteria not met
    - “Criteria to calculate annual mean not met.”
- More than 3 unique values (cite this)
    - 3 per year? 3 per quarter?
- What can we do for incomplete seasons or years: Compare to acute? Summarize based on number of values above detection?   (Cassie did this)


<br> __Annual reporting rules__


## Completeness checks  {#complete}

Methods for performing completeness checks...

Packages
```{r message=F}
library(tidyverse)

```

Our example data is organized by monitoring site and date. Here's a sample.

```{r kable, message=F, echo=F, fig.cap = "Sample data table."}
library(knitr)

df <- read_csv(
'"AQS_ID","Date","Conc"
270535501,"2009-07-30",0.00148
270535501,"2009-09-30",0.00064
270535501,"2009-11-30",0.34256
270535501,"2009-12-30",0.00064
270535502,"2009-03-30",0.26219
270535502,"2009-07-30",0.01113
270535502,"2009-09-30",0.00044
270535502,"2009-11-30",0.00127
270535502,"2009-12-30",0.00113')

kable(df)
```
